### 2024-11-25 paper/al.tex
Add 5.4

SpecTec introduces AL, an algorithmic specification language, for specifying
the prose specification.
Since the prose notation of the official specification is informal, it is
essential to rigorously define AL to ensure an unambiguous and correct
specification, rather than merely imitating the original.
We begin by outlining the ambiguity of the prose specification in
\cref{sec:prose-informal}.
Next, we propose a formal model for Wasm control flow in AL in
\cref{sec:model}.
We then define the syntax and the semantics of AL in \cref{sec:definition}.
\red{Finally, we ??? in \cref{???}.}

@@

\text{NOTE: $e_1$ is either store or current frame} \\

@@

\section{???}
\label{sec:???}

SpecTec translates declarative definitions in the DSL into algorithmic
definitions in AL.
\cref{ex:al-testop} is generated from \cref{ex:dsl-testop}, representing the
algorithmic specification for the \texttt{testop} instruction using AL:
\\
\begin{example}
\begin{lstlisting}[style=al]
@\textbf{Step\_pure/testop}@ nt testop {
    Assert (check_type_of_stack_top(nt))
    Pop (CONST nt c_1)
    Let c = @$\dollarcode$@testop_(nt, testop, c_1)
    Push (CONST I32 c)
}
\end{lstlisting}
  \label{ex:al-testop}
\end{example}
Using AL, SpecTec generates reStructuredText code.
\cref{fig:spectec-testop} shows the generated document for \texttt{testop}
instruction.
The generated document closely resembles the official document except for some
minor notation changes and some missing hyperlinks.

>>>>

SpecTec introduces AL, an algorithmic specification language, for specifying
the prose specification.
Since the prose notation of the official specification is informal, it is
essential to rigorously define AL to ensure an unambiguous and correct
specification, rather than merely imitating the original.
We begin by outlining the ambiguity of the prose specification in
\cref{sec:prose-informal}.
Next, we propose a formal model for Wasm control flow in AL in
\cref{sec:model}.
We then define the syntax and the semantics of AL in \cref{sec:definition}.
Finally, we produce an executable Wasm specification using AL interpreter that
ensures the correctness of the specification in \cref{sec:exec-spec}.

@@

@@

\section{An Executable WebAssembly Specification}
\label{sec:exec-spec}

% AL spec
SpecTec translates declarative definitions in the DSL into algorithms in AL.
\cref{ex:al-testop} is the generated algorithm from \cref{ex:dsl-testop},
representing the algorithmic specification for the \texttt{testop} instruction:
\\
\begin{example}
\begin{lstlisting}[style=al]
@\textbf{Step\_pure/testop}@ nt testop {
    Assert (check_type_of_stack_top(nt))
    Pop (CONST nt c_1)
    Let c = @$\dollarcode$@testop_(nt, testop, c_1)
    Push (CONST I32 c)
}
\end{lstlisting}
  \label{ex:al-testop}
\end{example}
SpecTec generates reStructuredText code from the AL algorithms.
The upper part of \cref{fig:spectec-testop} shows the rendered result of
reStructuredText code for the \texttt{testop} instruction.

% Executable
A notable feature of AL is executability, supported by SpecTec's AL interpreter
following the syntax and the semantics of AL.
This executability allows the specification itself to be executed, enabling the
execution of the specified language, Wasm.
The execution of a Wasm program begins with two entry points: module
instantiation and function invocation, both defined within the specification as
algorithms.
Executing these algorithms performs operations for executing the Wasm program,
ultimately producing its execution result.
In this way, the execution of a Wasm program is realized by the execution of
the specification itself.

% Correctness
That is to say, the prose specification written in AL, together with the AL
interpreter, functions as a Wasm interpreter which faithfully follows the
written specification.
By testing this Wasm interpreter, we can test the specification, thereby
ensuring its correctness.

### 2024-11-25 paper/intro.tex
Add 5.4

% Solution
To address this, we present a formal model for how AL describes Wasm control
flow.
Using this model, we identify the underlying issues and the necessary changes
to improve it.
Based on these insight, we formally define AL's syntax and semantics to
accurately capture Wasm control flow.
We then implement an AL interpreter according to this formalization and
evaluate its correctness.
The interpreter successfully passes all tests in the Wasm official test suite,
as well as tests from proposals.

% Contribution
The contributions of this paper are as follows:
\begin{itemize}
  \item We propose a formal model for Wasm control flow.
    (\cref{sec:model})
  \item We define syntax and semantics of AL (\cref{sec:definition})
  \item We implement an AL interpreter based on the formalization
    (\cref{ch:interpreter})
\end{itemize}

>>>>

% Solution
To address this, we present a formal model for how AL describes Wasm control
flow.
Using this model, we identify the underlying issues and the necessary changes
to improve it.
Based on these insight, we formally define AL's syntax and semantics to
accurately capture Wasm control flow.
We then implement an AL interpreter according to the formal definition and
evaluate its correctness.
The interpreter successfully passes all tests in the Wasm official test suite,
as well as tests from proposals.

% Contribution
The contributions of this paper are as follows:
\begin{itemize}
  \item We propose a formal model for Wasm control flow. (\cref{sec:model})
  \item We define syntax and semantics of AL (\cref{sec:definition})
  \item We implement an AL interpreter for executable specification
    (\cref{sec:exec-spec})
\end{itemize}
